---
title: "assign1-sta-1039580"
author: "Jiayu Wang"
date: "4/20/2021"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Q1a
```{r Q1a, echo= T}
#Read data
setwd("C:/Users/iefad/Desktop/course2021/bioinfosta/assignment/data-assignment1")
gt1 <- read.csv('genotypes1.csv')
gt2 <- read.csv('genotypes2.csv')
gt3 <- read.csv('genotypes3.csv')
pt1 <- read.csv('phenotype1.csv')
pt2 <- read.csv('phenotype2.csv')

#check the study sizes. As shown in the output below, the study sizes are consistent with what's reported.
dim(gt1) # sample size for genotypes1
dim(gt2) # sample size for genotypes2
dim(gt3) # sample size for genotypes3
dim(pt1) # sample size for phenotypes1
dim(pt2) # sample size for phenotypes2

#Show the first 5 phenotypes in study1 and study2.
pt1[1:5,2]  # in study 1
pt2[1:5,2]  # in study 2
```

## Q1b

```{r Q1b, echo=T}
recordpvalue <- data.frame(genotype = colnames(gt1)[2:201], pvalue = rep(0, 200))
for (i in 1:200) {
  model <-  lm(pt1$BMI ~ gt1[,i+1])
  recordpvalue[i,2] = anova(model)[1,5]
}
recordpvalue # record p-value 
```

## Q1d
```{r Q1d, echo=T}
#As shown in output, snp052 has the smallest p-value at 2.28028e-07.
recordpvalue[order(recordpvalue$pvalue), ][1,]
```
## Q2a
```{r Q2a, echo=T}
colnames(pt1)[2] = 'overweight'
pt1[,2] = as.logical(pt1[,2] > 25) # treat phenotypes from study1 the same as study2
pt = rbind(pt1,pt2)  #creating a single data frame for the phenotype
gt = rbind(gt1,gt2)  ##creating a single data frame for the genotype
```
## Q2b
```{r Q2b, echo=T}
combinepvalue <- data.frame(genotype = colnames(gt)[2:201], pvalue = rep(0, 200))
for (i in 1:200) {
  model <-  lm(as.numeric(pt$overweight) ~ as.numeric(gt[,i+1]))
  combinepvalue[i,2] = summary(model)$coefficients[2,4]
}
combinepvalue
```

## Q2d
```{r Q2d, echo=T}
#As shown in output, snp052 has the smallest pvalue at 2.853124e-14
sortedpvalue <- combinepvalue[order(combinepvalue$pvalue), ]
sortedpvalue[1,]
```
## Q2e(i)
```{r Q2ei, echo=T}
#The number is 16.
m = length(combinepvalue$pvalue)
sum(combinepvalue$pvalue < 0.05/m)
```
## Q2e(ii)
```{r Q2eii, echo=T}
#The number is 29.
max(which(sort(combinepvalue$pvalue) <= 0.05*seq(1, m)/m))
```
## Q2e(iii)
```{r Q2eiii, echo=T}
#The number is 21, the resulting FDR estimate is 0.007407407 as shown in output.
hist(combinepvalue$pvalue,n=20)
nsig = sum(combinepvalue$pvalue < 0.001)
nullfrac = mean(combinepvalue$pvalue>0.1)/0.9
falsep = nullfrac*m*0.001
FDR = falsep/nsig
print(list(nsig,FDR))
```
## Q2f
???????

## Q3a
```{r Q3a, echo=T}
#The SNPs with 8 smallest p-values and their p-values are shown as output.
combinepvalue[order(combinepvalue$pvalue),][1:8, ]
```
## Q3b
```{r Q3b, echo=T}
#The correlation matrix is shown in output.
#The linkage disequilibrium between some pairs of SNPs are high with correlation higher than 0.5.  
small8 <- combinepvalue[order(combinepvalue$pvalue),][1:8, ]
cormatrix <- matrix(ncol = 8, nrow = 8)

colnames(cormatrix) <- small8$genotype
rownames(cormatrix) <- small8$genotype

index <- as.numeric(row.names(small8))

for (i in 1:8) {
  for (j in 1:8) {
    cormatrix[i,j] <- cor(as.numeric(gt[,index[i]+1]),as.numeric(gt[,index[j]+1]))
  }
}
cormatrix
sum(cormatrix>0.5) #check the correlation > 0.5 which might be helpful to conclude
```
## Q3c
```{r Q3c, echo=T}
i <- 1
j <- 1
while (i <= 2){
  while (j <= nrow(cormatrix)){
    if (cormatrix[i,j] > 0.5 & (i != j)){
      cormatrix <- cormatrix[-j,-j]
    }
    else {
      j <- j+1
    }
  }
  j <- 1
  i <- i+1
}
cormatrix
```

left_index
index[!index %in% (index[3])]

## Q4a
```{r Q4a, echo=T}
#The parameter estimates and standard errors are shown in the output. 
overweightstate <- pt$overweight
has_052 <- as.numeric(gt[,53])
has_107 <- as.numeric(gt[,108])

fit1 = glm(overweightstate ~ has_052 + has_107,family="binomial")
coef(summary(fit1))[2:3,1:2]
```

#Q4b
The OR for SNP052 and SNP107 are 1.653549 and 1.495694 respectively.

Odds ratio = ???????????????
```{r Q4b, echo=T}
# Odds ratio = 
beta052 <- as.numeric(coef(fit1)[2])
beta107 <- as.numeric(coef(fit1)[3])

OR_052 <- exp(beta052)
OR_107 <- exp(beta107)

OR_052
OR_107
```
#Q4c
C.I for SNP052 is (1.425441, 1.918160).
C.I for SNP052 is (1.294375, 1.728325).
```{r Q4c, echo=T}
#Calculate SE
se_052 <- coef(summary(fit1))[2,2]
se_107 <- coef(summary(fit1))[3,2]
se_052 
se_107

#Calculate C.I.
C.I_052 <- exp(beta052 + c(-1,1) * qnorm(0.975) * se_052)
C.I_107 <- exp(beta107 + c(-1,1) * qnorm(0.975) * se_107)

C.I_052
C.I_107

```
#Q5a
The risk is as shown in the output.
```{r Q5a, echo=T}
new <- data.frame(has_052 = as.numeric(gt3[,53]), has_107 = as.numeric(gt3[,108]))
prerisk <- predict.glm(fit1,new,type = "response")
prerisk

```


#Q5b
The risk is as shown in the output.
```{r Q5b, echo=T}
premodel <- data.frame(individuals = gt3$X, risk = prerisk)
premodel[order(premodel$risk), ]
plot(premodel[order(premodel$risk), ]$risk, xlab = "sorted individuals", ylab = "risk", main = "risks with the individuals sorted in order of increasing risk")

```

#Q5c
The risk of indiv2001 is 0.580312.
```{r Q5c, echo=T}
premodel[premodel$individuals=='indiv2001',]

```

#Q5d
The OR is 4.089562.
```{r Q5d, echo=T}
maxrisk <- max(prerisk)
minrisk <- min(prerisk)

ODmax <- maxrisk/(1 - maxrisk)
ODmin <- minrisk/(1 - minrisk)
OR <- ODmax/ODmin
OR
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.



